---
title: "Biostat776 - Project1 WE"
author: "Wendy Camejo Nunez"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}
```

```{r}
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)
glimpse(chocolate)
```

## Part 1: Explore data

1.Make a histogram of the rating scores to visualize the overall distribution of scores. Change the number of bins from the default to 10, 15, 20, and 25. Pick on the one that you think looks the best. Explain what the difference is when you change the number of bins and explain why you picked the one you did.
```{r}
library(dplyr)
library(ggplot2)
chocolate %>%
    count(rating)
qplot(rating, data = chocolate)
qplot(rating, data = chocolate, bins=10)
qplot(rating, data = chocolate, bins=15)
qplot(rating, data = chocolate, bins=20)
qplot(rating, data = chocolate, bins=25)
```

I would choose 15 the other options make the I dont know

2.Consider the countries where the beans originated from. How many reviews come from each country of bean origin?
```{r}
bean_origin <- chocolate %>% count(country_of_bean_origin)
glimpse(bean_origin)
 ggplot(data=bean_origin,
       aes(x = `country_of_bean_origin`, y = `n`)) +
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle=90, vjust= 0.5, hjust=1))+
  ylab("Number_reviews") +xlab("Country_of_bean_origin")
```

3.What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? For this same set of reviews, also calculate (1) the total number of reviews and (2) the standard deviation of the rating scores. Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.
```{r}
dataEcuador <- chocolate %>% filter(country_of_bean_origin == "Ecuador")
glimpse(dataEcuador)

summarize(dataEcuador,
          mean= mean(dataEcuador$rating),
          sd= sd(dataEcuador$rating),
          total= length(dataEcuador$rating))


```
The average rating scores id 3.16.

4.Which company (name) makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?
```{r}
#Start by selecting columns 2,5,10 from the chocolate data
selectcolumns <- chocolate%>%select(2,5,10)%>%filter(country_of_bean_origin == "Ecuador")
glimpse(selectcolumns)
#change the order by company name
reorder_data <- arrange(selectcolumns, company_manufacturer)
glimpse(reorder_data)
data_1 <- reorder_data %>% group_by(company_manufacturer)%>% summarize(value=mean(rating))
data_1 <- arrange(data_1, desc(value))
glimpse(data_1)
head(data_1)
```
Amano, Benoit Nihant, Beschle (Felchlin), Durci, and Smooth Chocolator share the highest rating on average.

5.Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?
```{r}
top3 <- chocolate%>% group_by(country_of_bean_origin)%>% summarize(value=mean(rating))
glimpse(top3)
top3 <- arrange(top3, desc(value))
glimpse(top3)
```
Tobago, China, and Sao Tome & Principe have the highest rating on average.

6. Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?
```{r}
data_2 <- chocolate%>%select(5,10)
data_2 <- arrange(data_2, country_of_bean_origin)
data_2.2 <- filter(bean_origin, n>10)
glimpse(data_2.2)
c.data <- inner_join(data_2.2, data_2, by="country_of_bean_origin")
glimpse(c.data)

top3_1 <- c.data%>% group_by(country_of_bean_origin)%>% summarize(value=mean(rating))
glimpse(top3_1)
top3_1 <- arrange(top3_1, desc(value))
glimpse(top3_1)
```
Now the top 3 countries are Congo, Cuba, and Vietnam with more than 10 reviews.

7.For this last part, let’s explore the relationship between percent chocolate and ratings.
```{r}
more_50 <- filter(bean_origin, n>50)
glimpse(more_50)
more_50.1 <- inner_join(more_50,data_2, by="country_of_bean_origin")
glimpse(more_50.1)
```




## Part 2: Join two datasets together

1.Use this dataset it to create a new column called continent in our chocolate dataset that contains the continent name for each review where the country of bean origin is.
```{r}
#install.packages("gapminder")
library(gapminder)
library(dplyr)
glimpse(gapminder)
view(gapminder)

# deleting any duplicate countries from gapminder
my_data <- gapminder%>%distinct(country, .keep_all = TRUE)
view(my_data)

#selecting only country continent and country_of_bean_origin
gapminder1 <- my_data%>%select(1,2)

gapminder2 <- rename(gapminder1, country_of_bean_origin=country)

data3 <- left_join(x=chocolate, y=gapminder2, by="country_of_bean_origin")

data3_arrange <- arrange(data3, by=country_of_bean_origin)
view(data3)
view(data3_arrange)


# I need to change the NA to continent
```

2.Only keep reviews that have reviews from countries of bean origin with at least 10 reviews.
```{r}
country_1 <- data3_arrange%>%count(country_of_bean_origin)
country_2 <- filter(country_1, n>=10)

mas_10 <- inner_join(data3_arrange, country_2, by="country_of_bean_origin")
glimpse(mas_10)
#n=total number of reviews
```
3.Also, remove the country of bean origin named "Blend".
```{r}
data_4 <- subset(mas_10, country_of_bean_origin!="Blend")
glimpse(data_4)
```

4.Make a set of violin plots with ratings on the y-axis and continents on the x-axis.
```{r}
data_4%>%ggplot(aes(x=continent,y=rating)) + geom_violin()
```

## Part 3: Converting wide data into long data
1.Create a new set of columns titled beans, sugar, cocoa_butter, vanilla, letchin, and salt that contain a 1 or 0 representing whether or not that review for the chocolate bar contained that ingredient (1) or not (0).
```{r}
data_5 <- chocolate%>% mutate(beans = ifelse(grepl ("B", chocolate$ingredients), "1", "0"))
view(data_5)
data_5.1 <- data_5%>% mutate(sugar = ifelse(grepl ("S", chocolate$ingredients), "1", "0"))
data_5.2 <- data_5.1%>% mutate(cocoa_butter = ifelse(grepl ("C", chocolate$ingredients), "1", "0"))
data_5.3 <- data_5.2%>% mutate(vanilla = ifelse(grepl ("V", chocolate$ingredients), "1", "0"))
data_5.4 <- data_5.3%>% mutate(letchin = ifelse(grepl ("L", chocolate$ingredients), "1", "0"))
data_5.5 <- data_5.4%>% mutate(salt = ifelse(grepl ("S", chocolate$ingredients), "1", "0"))
glimpse(data_5.5)
view(data_5.5)

```

2.Create a new set of columns titled char_cocoa, char_sweet, char_nutty, char_creamy, char_roasty, char_earthy that contain a 1 or 0 representing whether or not that the most memorable characteristic for the chocolate bar had that word (1) or not (0). For example, if the word “sweet” appears in the most_memorable_characteristics, then record a 1, otherwise a 0 for that review in the char_sweet column (Hint: check out str_detect() from the stringr package).
```{r}
data_5.6 <- data_5.5%>% mutate(char_cocoa = ifelse(grepl ("cocoa", chocolate$most_memorable_characteristics), "1", "0"))
data_5.7 <- data_5.6%>% mutate(char_sweet = ifelse(grepl ("sweet", chocolate$most_memorable_characteristics), "1", "0"))
data_5.8 <- data_5.7%>% mutate(char_nutty = ifelse(grepl ("nutty", chocolate$most_memorable_characteristics), "1", "0"))
data_5.9 <- data_5.8%>% mutate(char_creamy = ifelse(grepl ("creamy", chocolate$most_memorable_characteristics), "1", "0"))
data_5.10 <- data_5.9%>% mutate(char_roasty = ifelse(grepl ("roasty", chocolate$most_memorable_characteristics), "1", "0"))
data_5.11 <- data_5.10%>% mutate(char_earthy = ifelse(grepl ("earthy", chocolate$most_memorable_characteristics), "1", "0"))
glimpse(data_5.11)
```

3.For each year (i.e. review_date), calculate the mean value in each new column you created across all reviews for that year. (Hint: If all has gone well thus far, you should have a dataset with 16 rows and 13 columns).
```{r}
data_5.11$beans = as.numeric(as.character(data_5.11$beans))
data_5.11$sugar = as.numeric(as.character(data_5.11$sugar))
data_5.11$cocoa_butter = as.numeric(as.character(data_5.11$cocoa_butter))
data_5.11$vanilla = as.numeric(as.character(data_5.11$vanilla))
data_5.11$letchin = as.numeric(as.character(data_5.11$letchin))
data_5.11$salt = as.numeric(as.character(data_5.11$salt))
data_5.11$char_cocoa = as.numeric(as.character(data_5.11$char_nutty))
data_5.11$char_creamy = as.numeric(as.character(data_5.11$char_creamy))
data_5.11$char_roasty = as.numeric(as.character(data_5.11$char_roasty))
data_5.11$char_earthy = as.numeric(as.character(data_5.11$char_earthy))
data_5.11$char_s = as.numeric(as.character(data_5.11$char_earthy))
glimpse(data_5.11)



data_7 <- data_5.11 %>%                   
  group_by(review_date) %>%
  summarise(beans_mean = mean(data_5.11$beans), sugar_mean = mean(data_5.11$sugar),cocoa_butter_mean = mean(data_5.11$cocoa_butter),vanilla_mean = mean(data_5.11$vanilla),letchin_mean = mean(data_5.11$letchin),salt_mean = mean(data_5.11$salt),char_cocoa_mean = mean(data_5.11$char_cocoa),char_nutty_mean = mean(data_5.11$char_nutty),char_creamy_mean = mean(data_5.11$char_creamy),char_roasty_mean = mean(data_5.11$char_roasty),char_earthy_mean = mean(data_5.11$char_earthy),char_sweet_mean = mean(data_5.11$char_sweet),.groups = 'drop') %>% as.data.frame
glimpse(data_7) 
```
4.Convert this wide dataset into a long dataset with a new feature and mean_score column.
```{r}

```

## Part 4: Data visualization
Use the functions in ggplot2 package to make a scatter plot of the mean_scores (y-axis) over time (x-axis). One point for each mean_score. For full credit, your plot should include:
```{r}

```
## Part 5: Make the worst plot you can!
Using the chocolate dataset (or any of the modified versions you made throughout this assignment or anything else you wish you build upon it):
```{r}
 ggplot(data=bean_origin,
       aes(x = `country_of_bean_origin`, y = `n`)) +
  theme(axis.text.x = element_text(angle=180, vjust= 5, hjust=5), legend.position = "top")+
labs(title = "The begining of the World", x="x", y="y")+ geom_boxplot()+ geom_point(aes(colour = "red")) + geom_smooth()
  
```
-I choose the ggplot that I made in question 1.2 showing country_of_bean_origin and Number_reviews.
- I changed the angle, vjust and hjust for the x-axis which makes it impossible to read its content.
- I added a title that it is unrelated to content of the plot using the function labs().
- I changed the names of the x- and y-axis for "x" and "y" using the function lab() which doesn't provide any information.
- I changed the plot from a geom_bar() to geom_boxplot which makes each of the points to have bar coming out of it.
- I changed the color of the point using geom_point() with astetic funtion from balck to red
_ I changed the position of the legend from the rigth side to the top.

## Part 6: Make my plot a better plot
The goal is to take my sad looking plot and make it better! If you’d like an example, here is a tweet I came across of someone who gave a talk about how to zhoosh up your ggplots.
```{r}
chocolate %>%
    ggplot(aes(
        x = as.factor(review_date),
        y = rating,
        fill = review_date)) + 
theme(axis.text.x = element_text(angle = 90,size =12, colour = "black", vjust = 0.5, hjust = 1, face = "bold"), 
axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 12, face = "bold"), 
        legend.text = element_text(size = 12, face = "bold", colour = "dark blue"), 
        axis.text.y = element_text(colour = "black", size = 12, face = "bold")) +
    geom_violin()+ geom_boxplot(width= .1, fill= "gray") +labs(title = "Chocolate Review Date",subtitle = "2020 and 2021 have the overall highest reviews since 2006", x= "Review Date", y="Rating")

```
- Added a title to the graph using the function labs(title)
- Added a subtitle to give the main results of the graph using the function labs(subtitle)
- Change the x-axis title form "as.factor(review_date)" to "Review Date" 
- Change the axis and legend font to size to 12 and made it bold to be more clear for readers using the function theme().
- Change the color of the numbers inside of the legend to make the graph more appealing to readers using the function legend.text
- Change the date in x-axis from horizontal position to vertical position to have more space between each other using element_text(angle = 90)inside of the theme function.
- Added geom_boxplot() in color gray to help visualize the median of the plot.


